Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems

### 2020-05-10 manuscript/bef_stability/appendix.Rmd
Commit before text review

```{r, echo = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

@@

```{r}
caption_indir_stab_sem <- paste0(
  "Indirect and total effect of variables on biomass stability from the
  structural equation model. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.61$. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "Network structure being the indirect effect that went through both avg trophic
  level and connectance and Species richness_Network structure being the
  indirect effect that went through (i) species richness and from (ii) species
  richness to network structure."

caption_indir_bm_sem <- paste0(
  "Indirect and total effect of variables on total biomass from the
  structural equation model. See the table (X) for details about computation and
  meaning."

)
```

```{r indir-tab, fig.cap = "Indirect effect from SEM"}
format_table <- function (x = tmp_stab_indir_table) {
  x <- x[, c("variable", names(x)[!names(x) %in% "variable"])]

  rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
  names(rm_n) <- names(get_sem_var_name_replacement()) 
  x$variable <- str_replace_all(x$variable, rm_n)

  colnames(x) %<>% str_remove_all(., pattern = "via_")

  col_replacement <- c(
    variable = "Variable",
    cv_sp = "CVsp",
    sync = "Synchrony",
    richness = "Species richness",
    com = "Network structure",
    richness_com = "Sp. richness and network str.",
    total = "Total effect"
  )
  colnames(x) %<>% str_replace_all(., col_replacement)

  x
}

kable(format_table(tmp_stab_indir_table),
  "latex",
  booktabs = T,
  caption = caption_indir_stab_sem 
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))

kable(format_table(tmp_bm_indir_table),
  "latex",
  booktabs = T,
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

>>>>

```{r, echo = FALSE, message = FALSE, collapse = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

@@

### 2020-05-10 manuscript/bef_stability/manuscript.Rmd
Commit before text review

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)

@@

#### Importance of BEF assessment

Determining the shape of the relationship between species diversity and ecosystem
functioning has been a central research topic in ecology
[@may_will_1972, @hector_plant_1999, @loreau_biodiversity_2001, @hooper_effects_2005, @duffy_why_2009, @isbell_forest_quantifying_2018].
Current biodiversity losses
in all ecosystems challenge scientists to understand how ecosystems will be
affected in turn. Studies in Bioversity-Ecosystem Functioning (BEF) usually
quantify species diversity as species richness (i.e number of species) while
ecosystem functioning is quantified as total biomass, productivity (biomass
produced by unit of time) or stability which is quantified in more diverse ways
such as the variability of species abundance, the resilience to perturbation, or
resistance to perturbation [@donohue_navigating_2016].

@@

Species richness has been reported to increase productivity
[@tilman_productivity_1996, @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partitioning_2001, @cardinale_species_2002, @fargione_selection_2007].
Empirical assessment of the effect of species richness on total biomass showed negative
[@grace_integrative_2016] and positive relationships
[@hector_plant_1999, @tilman_biodiversity_2014]. @grace_integrative_2016 showed
that the results depend on the refinement of the mecanisms involved in the
analysis.

@@

The relationship between species richness and stability is even more complex
partly because of the inherent multidimensionality of stability
[@donohue_dimensionality_2013, @donohue_navigating_2016] and because of the
different metrics of stability privileged in theoretical (stability of local
equilibrium, robustness to primary extinction) and empirical work (variability
of abundance or biomass through time). Empirical studies showed consistently
that species richness decreased community variability in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018] and
vertebrates [@olivier_independent_nodate, @franssen_annual_2011]. Two
mecanisms are involved into community variability, (1) species
variability and (2) the species synchrony [@loreau_species_2008, @thibaut_understanding_2013].
Positive [@tilman_biodiversity_2006], neutral
[@jiang_different_2009, @olivier_independent_nodate] and negative
[@franssen_annual_2011, @olivier_independent_nodate] relationships between
species variability and species richness have been found. On the other side,
studies consistently found that species richness decreased synchrony between
species
[@loreau_species_2008, @thibaut_understanding_2013, @bluthgen_land_2016, @olivier_independent_nodate].
Beyond species richness, species interactions, which modifies species
abondance and biomass and make species dynamics interdependant, have been
largely ignored in empirical assessment.

@@

Conversely, theoretical ecology display an extensive literature about the
effects of network structure on stability 
[@may_will_1972, @angelis_stability_1975, @stouffer_compartmentalization_2011, @thebault_stability_2010, @barbier_pyramids_2019, @shanafelt_david_w._stability_nodate].
Theoretical ecology generally uses stability definitions related to community
equilibrium and its stability to small perturbations.
Connectance (i.e. proportion of realized links in a network) can decrease
[@angelis_stability_1975, @thebault_stability_2010] or increase resilience of
trophic network depending on the functioning of networks
[@angelis_stability_1975].
The number of trophic levels, the vertical position in the trophic network as
well as the strength of their biomass regulation (i.e. top-down and bottom-up)
can determine biomass and stability at equilibrium
[@hurd_stability_1971, @shanafelt_david_w._stability_nodate, @zhao_horizontal_2019, @barbier_pyramids_2019].

@@

@@

```{r}
matmet_cap <- paste0(
  "Mat&met figure"
)
```

```{r matmet, fig.cap = matmet_cap}
myload(p_st,temporal_network, dir = mypath("manuscript/bef_stability/figs"))
p_st <- p_st +
  theme(legend.position = "none")
# Network inference
net_method_path <- mypath("manuscript/bef_stability/figs",
  "network_inference.pdf")
net_method <- net_method_path %>%
  magick::image_read()
p_net_method <- cowplot::ggdraw() +
  cowplot::draw_image(net_method)

top <- plot_grid(p_st, p_net_method, ncol = 2, rel_widths = c(.8, 1), labels =
  "AUTO")
p_method <- plot_grid(top, temporal_network, nrow = 2, rel_heights = c(1, .3),
  labels = c("", "C"))
p_method
```

@@

To obtain the network structure of the communities, we needed to infer
ecological interactions between the components of the communities and thus build
the ecological network. As the interactions were not measured, we used trait
matching method to infer the presence or absence of an interaction. This method is
widely used to infer trophic interactions
[@brose_predator_2019, @poisot_structure_2016, @schneider_body_2012]. Indeed,
the existence of an trophic interaction between two species have been reported
to be largely dependant on their body size ratio [cascade model, @cohen_stochastic_1985, @warren_invertebrate_1987, @brose_predator_2019].

@@

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies
[@brose_predator_2019, @poisot_structure_2016, @schneider_body_2012].
For a given trophic species, we first determined if its size class end up in a
life stage containing piscivorous diet, i.e. if there was an overlap between
them. We were more liberal for piscivory determination to not miss potential
fish-fish trophic interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size range
from 3% to 45% of the median of its size range. A trophic interaction was set
for each trophic species whose median of its size class fell into the prey range
of the predator trophic species. To sum up, the metaweb was obtained with
allometric feeding interactions (1) between trophic species, (2) between trophic
species and resource nodes and (3) between resource nodes.

@@

Habitat description were obtained from fish sampling data. We kept variables
describing the ecosystem size namely the stream width, depth and the ecosystem
position (altitude and slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature as physical variables to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018, @tonkin_flow_2018, @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water.
The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects and took the annual mean (i.e. one
value by year and by station. As the environmental recording stations were not
located at the same place than the fish stations, we interpolated environmental
data with spatial statistical network model [@isaak_applications_2014]
implemented in the `SSN` R package [@hoef_ssn:_2014]. The interpolation
procedure is detailed in supplementary material. For each fish station,
environmental variables were matched with the year of fish sampling (except for
water temperature for which the record began in 2006). Then, the average
temperature, flow, BOD, river width and depth were computed as well as
coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and environmental
variables are often colinear, a principal component analysis (PCA) was performed
on the scaled variables. The PCA explained `r tot_var_expl_pca`% of the
variance. An axis rotation was performed to maximize the representation of the
variables by the principal components (Fig. \@ref(fig:pca-rotated)). We kept
five axis, representing respectively the (1) river size (river width & depth),
(2) the altitude, slope and decreasing average temperature, (3) average
enrichment (BOD) and CV of flow, (4) river size CV (CV of river width and depth)
and (5) CV of enrichment (CV of BOD). Additionally, we logged the axis one, two
and three to normalize their distribution and we reversed the third axis to get
increasing average temperature with increasing values of the axis.

@@

We seeked to disantangle the unravel the links between
environment and community structure on both biomass stability and total biomass.
To investigate complex relationships, direct and indirect, between variables,
Structural Equation Modeling (SEM) is one promising way  
[@grace_structural_2008, @grace_causal_2014, @lefcheck_piecewisesem_2016,
@grace_integrative_2016]. SEM makes possible to set causal relationships between
variables. This step can be critical, as the estimated parameters can be
determined by both the choice of the variables and the links between them.
Therefore, the set of variables and links shoud be supported by theory and
litterature.

@@

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

@@

```{r get-sem-tot-effect-env}
stab_all_params <- get_coefficient_piecewisesem(sem = stab_sem_rich, p_val_thl = NULL)
# effect of env on sync and cv_sp
env_on_sync <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
env_on_cv_sp <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
# env on stab via network 
env_on_t_lvl <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "t_lvl", type = "std")
env_on_ct <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "ct", type = "std")
env_on_richness <- get_env_on_rich(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
```

@@

```{r}
# effect of env on stab
env_on_stab <- get_env_on_stab(fit = stab_all_params, p_val_thl = 0.05,
  type = "std")
rich_on_stab <- get_rich_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
com_on_stab <- get_com_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, com_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()

options(knitr.table.format = "latex")
kable(tmp_stab_indir_table, "latex", booktabs = T, align = "c") 
kable(tmp_bm_indir_table, "latex", booktabs = T, align = "c")

#kable(rbind(
  #mutate(tmp_stab_indir_table, type = "stab"),
  #mutate(tmp_bm_indir_table, type = "bm")
  #), "latex", booktabs = T, align = "c") %>%
  #column_spec(1, bold=T) %>%
  #collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")
```

@@

We found a negative relationship biomass stability and species
richness whereas there was a positive one between total biomass and species
richness. Overall, it contrasts with studies in grasslands
[@tilman_biodiversity_2006, @zhang_scale_nodate], mesocosms
[@pennekamp_biodiversity_2018] or other animals [@olivier_independent_nodate]
where positive relationships are typically found between biomass stability
and species richness. In streams as well, @franssen_annual_2011
found that species richness increased biomass stability. However other studies
found a opposite relationship [@yang_effects_2011] or a marginal one
[@declerck_species_2006]. We found that species richness increased total biomass
as found in salt marshes, forests
[@wu_relationship_2015,@li_relationship_2018]. The positive effect
of species richness on total biomass might be due to resource use
complementarity [@cardinale_species_2002]. But opposite relation has been found
in richness in grasslands
[@grace_integrative_2016].

#### connectance on stability and biomass
Our path analysis revealed that, although species richness has a strong effect on
connectance, connectance itself had little to no effect either on stability nor
total biomass. This result contradicts precedent theoretical studies on the
relationship between stability and network structure [@angelis_stability_1975,
@may_will_1972], although they considered different stability metrics (i.e.
local stability of equilibrium point). The reason might be that in our inferred
network, network structure is strongly constraints by assembly processes.
Indeed, connectance generally decreased with increasing species richness
[@winemiller_must_1989, @thebault_stability_2010]. It is an effect important to
consider the link between connectance and ecosystem properties, as this direct
relationship can be misleading if species richness is not taken into account.

>>>>

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)

@@

#### Biodiversity and ecosystem functioning

Biodiversity is essential to determine and maintain ecosytem properties.
Current biodiversity losses question the ability of how ecosystems will be
affected in turn.

The relationship between biodiversity and ecosystem functioning has been a
central research topic in ecology [@may_will_1972; @hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @duffy_why_2009;
@isbell_forest_quantifying_2018].

Studies in
Bioversity-Ecosystem
Functioning (BEF) usually quantify species diversity as species richness (i.e
number of species) while ecosystem functioning is quantified as total biomass,
productivity (biomass produced by unit of time) or stability which is quantified
in more diverse ways such as the variability of species abundance, the
resilience to perturbation, or resistance to perturbation
[@donohue_navigating_2016].

@@

Species richness has been reported to increase productivity
[@tilman_productivity_1996; @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partitioning_2001; @cardinale_species_2002; @fargione_selection_2007].
Empirical assessment of the effect of species richness on total biomass showed
negative [@grace_integrative_2016] and positive relationships
[@hector_plant_1999; @tilman_biodiversity_2014]. @grace_integrative_2016 showed
that the results depend on the refinement of the mecanisms involved in the
analysis.

@@

The relationship between species richness and stability is even more complex
partly because of the inherent multidimensionality of stability
[@donohue_dimensionality_2013; @donohue_navigating_2016] and because of the
different metrics of stability privileged in theoretical (stability of local
equilibrium, robustness to primary extinction) and empirical work (variability
of abundance or biomass through time). Empirical studies showed consistently
that species richness decreased community variability in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018] and
vertebrates [@olivier_independent_nodate; @franssen_annual_2011]. Two mecanisms
are involved into community variability, (1) species variability and (2) the
species synchrony [@loreau_species_2008; @thibaut_understanding_2013].  Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. On the other side, studies consistently found
that species richness decreased synchrony between species [@loreau_species_2008;
@thibaut_understanding_2013; @bluthgen_land_2016; @olivier_independent_nodate].
Beyond species richness, species interactions, which modifies species abondance
and biomass and make species dynamics interdependant, have been largely ignored
in empirical assessment.

@@

Conversely, theoretical ecology display an extensive literature about the
effects of network structure on stability [@may_will_1972;
@angelis_stability_1975; @stouffer_compartmentalization_2011;
@thebault_stability_2010; @barbier_pyramids_2019;
@shanafelt_david_w._stability_nodate].  Theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small perturbations.  Connectance (i.e. proportion of realized links in a
network) can decrease [@angelis_stability_1975; @thebault_stability_2010] or
increase resilience of trophic network depending on the functioning of networks
[@angelis_stability_1975].  The number of trophic levels, the vertical position
in the trophic network as well as the strength of their biomass regulation (i.e.
top-down and bottom-up) can determine biomass and stability at equilibrium
[@hurd_stability_1971; @shanafelt_david_w._stability_nodate;
@zhao_horizontal_2019; @barbier_pyramids_2019].

@@

- Q1: diversity-stability, diversity-biomass in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3:

@@

```{r}
matmet_cap <- paste0(
  "(A) Conceptual model of the hypothesized relationships between environment,
  species richness, network metrics and total biomass / biomass stability. ",
  "Solid and dashed arrows represent respectively direct and indirect effects.
  ",
  "The blue, green and blue/green dashed arrows being respectively indirect
  effects going through species richness, network metrics and through both
  species richness and network metrics.",
  "(B) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. ", "Grey shades highlighted elevalation variation. ",
  "Inner borders draw the limits of the hydrographic basins. ",
  "(C) The metaweb. Each trophic species (N = 412) is represented by a node and
  each color represent a species (S = ). Resource nodes (7) were drawn in grey. ",
  "(D) Subset over network variation through time."
)
```

```{r matmet, fig.cap = matmet_cap}
myload(p_st,temporal_network, dir = mypath("manuscript/bef_stability/figs"))

# Network inference
net_method_path <- mypath("manuscript/bef_stability/figs",
  "hypothesis.pdf")
metaweb_path <- mypath("manuscript/bef_stability/figs",
  "metaweb.png")
library(magick)

p_net_method <- cowplot::ggdraw() +
  cowplot::draw_image(magick::image_read_pdf(net_method_path))
p_metaweb <- cowplot::ggdraw() +
  cowplot::draw_image(magick::image_read(metaweb_path))

top <- plot_grid(p_net_method, p_st, ncol = 2, rel_widths = c(1, 1), labels =
  "AUTO")
bottom <- plot_grid(p_metaweb, temporal_network, ncol = 2,
  rel_widths = c(1, 1), labels = c("C", "D"))

p_method <- plot_grid(top, bottom, nrow = 2, rel_heights = c(1, 1),
  labels = c("", "C"))
save_plot(
  filename = mypath("manuscript/bef_stability/figs",
    "method_fig.pdf"),
  p_method, nrow = 2, ncol =2,
  base_width = 7/2.54,
  base_asp = 0.3
)
p_method
```

@@

To obtain the network structure of the communities, we needed to infer
ecological interactions between the components of the communities and thus build
the ecological network. As the interactions were not measured, we used trait
matching method to infer the presence or absence of an interaction. This method is
widely used to infer trophic interactions
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. Indeed;
the existence of an trophic interaction between two species have been reported
to be largely dependant on their body size ratio
[cascade model; @cohen_stochastic_1985; @warren_invertebrate_1987; @brose_predator_2019].

@@

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012].
For a given trophic species, we first determined if its size class end up in a
life stage containing piscivorous diet, i.e. if there was an overlap between
them. We were more liberal for piscivory determination to not miss potential
fish-fish trophic interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size range
from 3% to 45% of the median of its size range. A trophic interaction was set
for each trophic species whose median of its size class fell into the prey range
of the predator trophic species. To sum up, the metaweb was obtained with
allometric feeding interactions (1) between trophic species, (2) between trophic
species and resource nodes and (3) between resource nodes.

@@

Habitat description were obtained from fish sampling data. We kept variables
describing the ecosystem size namely the stream width, depth and the ecosystem
position (altitude and slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature as physical variables to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018; @tonkin_flow_2018; @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water.
The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects and took the annual mean (i.e. one
value by year and by station. As the environmental recording stations were not
located at the same place than the fish stations, we interpolated environmental
data with spatial statistical network model [@isaak_applications_2014]
implemented in the `SSN` R package [@hoef_ssn:_2014]. The interpolation
procedure is detailed in supplementary material. For each fish station,
environmental variables were matched with the year of fish sampling (except for
water temperature for which the record began in 2006). Then, the average
temperature, flow, BOD, river width and depth were computed as well as
coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and environmental
variables are often colinear, a principal component analysis (PCA) was performed
on the scaled variables. The PCA explained `r tot_var_expl_pca`% of the
variance. An axis rotation was performed to maximize the representation of the
variables by the principal components (Fig. \@ref(fig:pca-rotated)). We kept
five axis, representing respectively the (1) river size (river width & depth),
(2) the altitude, slope and decreasing average temperature, (3) average
enrichment (BOD) and CV of flow, (4) river size CV (CV of river width and depth)
and (5) CV of enrichment (CV of BOD). Additionally, we logged the axis one, two
and three to normalize their distribution and we reversed the third axis to get
increasing average temperature with increasing values of the axis.

@@

We seeked to disantangle the unravel the links between
environment and community structure on both biomass stability and total biomass.
To investigate complex relationships, direct and indirect, between variables,
Structural Equation Modeling (SEM) is one promising way  
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. SEM makes possible to set causal relationships between
variables. This step can be critical, as the estimated parameters can be
determined by both the choice of the variables and the links between them.
Therefore, the set of variables and links shoud be supported by theory and
litterature.

@@

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```
```{r}
caption_indir_stab_sem <- paste0(
  "Indirect and total effect of variables on biomass stability from the
  structural equation model. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.61$. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "Network structure being the indirect effect that went through both avg trophic
  level and connectance and Species richness_Network structure being the
  indirect effect that went through (i) species richness and from (ii) species
  richness to network structure."

)

caption_indir_bm_sem <- paste0(
  "Indirect and total effect of variables on total biomass from the
  structural equation model. See the table (X) for details about computation and
  meaning."

```{r indir-bm, fig.cap = "Indirect effect from SEM"}
format_table <- function (x = tmp_stab_indir_table) {
  x <- x[, c("variable", names(x)[!names(x) %in% "variable"])]

  rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
  names(rm_n) <- names(get_sem_var_name_replacement()) 
  x$variable <- str_replace_all(x$variable, rm_n)

  colnames(x) %<>% str_remove_all(., pattern = "via_")

  col_replacement <- c(
    variable = "Variable",
    cv_sp = "CVsp",
    sync = "Synchrony",
    richness = "Horizontal diversity",
    com = "Vertical diversity",
    richness_com = "Horizontal & vertical diversity",
    total = "Total effect"
  )
  colnames(x) %<>% str_replace_all(., col_replacement)

  x
}

kable(format_table(tmp_bm_indir_table),
  "latex",
  booktabs = T,
  caption = caption_indir_bm_sem
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))

```{r}
stab_all_params <- get_coefficient_piecewisesem(sem = stab_sem_rich, p_val_thl = NULL)
# effect of env on stab
env_on_stab <- get_env_on_stab(fit = stab_all_params, p_val_thl = 0.05,
  type = "std")
rich_on_stab <- get_rich_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
com_on_stab <- get_com_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, com_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

```{r indir-stab}
kable(format_table(tmp_stab_indir_table),
  "latex",
  booktabs = T,
  caption = caption_indir_stab_sem 
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

@@

```{r get-sem-tot-effect-env}
# effect of env on sync and cv_sp
env_on_sync <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
env_on_cv_sp <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
# env on stab via network 
env_on_t_lvl <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "t_lvl", type = "std")
env_on_ct <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "ct", type = "std")
env_on_richness <- get_env_on_rich(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
```

@@

@@

We found a negative relationship biomass stability and species
richness whereas there was a positive one between total biomass and species
richness. Overall, it contrasts with studies in grasslands
[@tilman_biodiversity_2006; @zhang_scale_nodate]; mesocosms
[@pennekamp_biodiversity_2018] or other animals [@olivier_independent_nodate]
where positive relationships are typically found between biomass stability
and species richness. In streams as well, @franssen_annual_2011
found that species richness increased biomass stability. However other studies
found a opposite relationship [@yang_effects_2011] or a marginal one
[@declerck_species_2006]. We found that species richness increased total biomass
as found in salt marshes, forests
[@wu_relationship_2015;@li_relationship_2018]. The positive effect
of species richness on total biomass might be due to resource use
complementarity [@cardinale_species_2002]. But opposite relation has been found
in richness in grasslands
[@grace_integrative_2016].

#### connectance on stability and biomass
Our path analysis revealed that, although species richness has a strong effect on
connectance, connectance itself had little to no effect either on stability nor
total biomass. This result contradicts precedent theoretical studies on the
relationship between stability and network structure [@angelis_stability_1975;
@may_will_1972], although they considered different stability metrics (i.e.
local stability of equilibrium point). The reason might be that in our inferred
network, network structure is strongly constraints by assembly processes.
Indeed, connectance generally decreased with increasing species richness
[@winemiller_must_1989; @thebault_stability_2010]. It is an effect important to
consider the link between connectance and ecosystem properties, as this direct
relationship can be misleading if species richness is not taken into account.
